---
title: "Population genetics analyses compiled"
author: "Julia Harenčár"
date: "5/2/2022"
format: 
  html:
    code-fold: true
    code-tools: true
    code-overflow: wrap
    toc: true
    number-sections: true
theme:
  light: minty
  dark: superhero
editor_options: 
  chunk_output_type: console
---
### Load packages
```{r}
#| echo: false
library(ggplot2)
library(ggrepel)
library(patchwork)
library(dplyr)
library(tidyverse)
library(readr)
theme_set(theme_bw())
library(cowplot)
```

author: "Julia Harenčár" date: "5/2/2022" output: html_document
### setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 3,
	fig.path = "ANGSD/PCAngsd_figs/",
	fig.width = 5
)
knitr::opts_knit$set(root.dir = '/Users/Julia/Documents/GitHub/Population_genetics_vill_alle')
#knitr::opts_knit$set(root.dir = '/Users/Julia/Library/CloudStorage/GoogleDrive-jharenca@ucsc.edu/My Drive/GitHub/seq_analysis/ANGSD')
```
## PCAs
Below, I summarize the PCAs I generated by running PCAngsd on the genotype likelihood (beagle) file generated from clean bams with greater than 0.05x coverage WITH -sites specified based on bcftools site filtering.

#### PCA of LGAHVB

generally a 0.15 cutoff

```{r full_angsd1_AVH_PCA, echo=FALSE}
setwd("/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/")
LGAHVBpop <- read.table("ANGSD/angsd_out/LGAHVB_231220.pop.info")
cov <- as.matrix(read.table("ANGSD/angsd_out/PCAngsd/LGAHVB_231220.PCAngsd.cov"))
e <- eigen(cov)

PC1 <- e$vectors[,1]
PC2 <- e$vectors[,2]
spp <- LGAHVBpop[,2]
ID <- LGAHVBpop[,1]
PC12.LGAVHB.co0.15 <- data.frame(PC1, PC2, ID, spp)

# Calculate the proportion of variance explained by each PC
variance_explained <- e$values / sum(e$values)
variance_explained[1] # 0.115
variance_explained[2] # 0.101

ggplot(PC12.LGAVHB.co0.15, aes(x=PC1, y=PC2, fill=spp)) +
  geom_point(size = 3, pch = 21, colour="black") +
  scale_x_reverse() +
  xlab("PC1 (11.5%)") +
  ylab("PC2 (10.1%)") +
  theme_bw() + 
  annotate("text", x = -0.2, y = 0.075, label = paste("# sites = 7,195"), hjust = 1.1, vjust = -1.5, size = 4, color = "black")
  
#geom_text_repel(aes(label=ID), size =1.5, nudge_y = 0.03, max.overlaps = 30)
#ggtitle("AVH1.selected.sites.co0.15x") +
#geom_text_repel(aes(label=ID), size =3, nudge_y = 0.03)

  
## example of how to add repelled text labels:
#library(ggrepel)
#ggplot(mtcars, aes(wt, mpg, label = rownames(mtcars))) +
#  geom_text_repel() +
#  geom_point(color = 'red') +
#  theme_classic(base_size = 16)
```

#### PCAs of alle, vill, and hybrids only (ANGSD_1; only selected sites with 0.05x coverage cutoff)

PCAngsd run on beagle file from angsd1 (samtools; see 'ANGSD run tracking - full dataset' in `[g-drive]/GitHub/seq_analysis/ANGSD/angsd_runs` with estimation of independent allele frequencies. Plot created from output of the following code:\
`python pcangsd.py -beagle $IN_DIR/selected_sites_0.05co_AVH.genolike.beagle.gz -o $OUT_DIR/sites_0.15co_AVH_2_PCAngsd`

```{r full_angsd1_AVH_PCA, echo=FALSE}
#latest (260209)
setwd("/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/")
AVH <- read.table("ANGSD/angsd_out/AHV_260209.pop.info")
cov <- as.matrix(read.table("ANGSD/angsd_out/PCAngsd/AHV_260209.PCAngsd.cov"))
e <- eigen(cov)

PC1 <- e$vectors[,1]
PC2 <- e$vectors[,2]
spp <- AVH[,2]
ID <- AVH[,1]
PC12.AVH <- data.frame(PC1, PC2, ID, spp)

# remove low breadth individual that snuck in
PC12.AVH <- PC12.AVH[-46,]

# Calculate the proportion of variance explained by each PC
variance_explained <- e$values / sum(e$values)
variance_explained[1]

#number of sites in the dataset:
#num_sites <- "47,820"

# With hand pollinated likely backcrosses (vill pollen, hybrid mom)
ggplot(PC12.AVH, aes(x=PC1, y=PC2, fill=spp)) +
  geom_point(size = 3, pch = 21, colour="black") +
  scale_fill_manual(values=c("#CC79A7", "#56B4E9", "#FFD700"), name=paste("Species"), labels=c(expression(italic("C. allenii")),"hybrid", expression(italic("C. villosissimus")))) +
  scale_x_reverse() +
  theme_bw() +
  theme(legend.text.align = 0,legend.position = "top") + 
  xlab("PC1 ()") +
  ylab("PC2 (1.9%") + 
  labs(x = paste0("PC1 (", round(variance_explained[1], 4)*100, "%)"), y = paste0("PC2 (", round(variance_explained[2], 3)*100, "%)"))
  #geom_text_repel(aes(label=ID), size =1.5, nudge_y = 0.03, max.overlaps = 30)

ggsave("ANGSD/PCAngsd_figs/AHV_PCA_260209.png", device = "png", width = 10, height = 7, units = "cm",dpi = 600)
# NOTE, 19.561 and 19.589 likely with some ancestry from laev
################################################################################
#latest (240121)
#Number of sites after MAF filtering (0.05): 13227
AVHF1pop <- read.table("ANGSD/angsd_out/AHVF1_0.15co_240109.clean.pop.info")
cov <- as.matrix(read.table("ANGSD/angsd_out/PCAngsd/AHVF1_240109.PCAngsd.cov"))
e <- eigen(cov)

PC1 <- e$vectors[,1]
PC2 <- e$vectors[,2]
spp <- AVHF1pop[,2]
ID <- AVHF1pop[,1]
PC12.AVHF1.co0.15 <- data.frame(PC1, PC2, ID, spp)

# Calculate the proportion of variance explained by each PC
variance_explained <- e$values / sum(e$values)
variance_explained[2] # only like 2%, so very little and can ignore

#number of sites in the dataset:
num_sites <- 13227

# With hand pollinated likely backcrosses (vill pollen, hybrid mom)
ggplot(PC12.AVHF1.co0.15, aes(x=PC1, y=PC2, fill=spp)) +
  geom_point(size = 3, pch = 21, colour="black") +
  scale_fill_manual(values=c("#CC79A7", "#93c572","#56B4E9", "#FFD700"), name=paste("Species"), labels=c(expression(italic("C. allenii")), "F1","hybrid", expression(italic("C. villosissimus")))) +
  scale_x_reverse() +
  theme_bw() +
  theme(legend.text.align = 0) + 
  xlab("PC1 ()") +
  ylab("PC2 (1.9%") + 
  labs(x = paste0("PC1 (", round(variance_explained[1], 4)*100, "%)"), y = paste0("PC2 (", round(variance_explained[2], 3)*100, "%)")) +
   annotate("text", x = -0.17, y = 0.25, label = paste("# sites =", num_sites), hjust = 1.1, vjust = -1.5, size = 4, color = "black")
  
# Without hand pollinated likely backcrosses (vill pollen, hybrid mom) #other yellow: #F0E442
n_alle <- PC12.AVHF1.co0.15 %>% filter(spp == "ALLE") %>% nrow()
n_vill <- PC12.AVHF1.co0.15 %>% filter(spp == "VILL") %>% nrow()
n_hyb <- PC12.AVHF1.co0.15 %>% filter(spp == "HYB") %>% nrow()
  
AHV_PCA <- PC12.AVHF1.co0.15 %>% 
  filter(spp !="F1") %>% 
  ggplot(aes(x=PC1, y=PC2, fill=spp)) +
  geom_point(size = 3, pch = 21, colour="black") +
  scale_fill_manual(values=c("#CC79A7", "#56B4E9", "#FFD700"), name=paste("Species"), labels=c(bquote(italic("C. allenii")~"(n="*.(n_alle)*")"), bquote(~"hybrid"~"(n="*.(n_hyb)*")"), bquote(italic("C. villosissimus")~"(n="*.(n_vill)*")"))) +
  theme_bw() +
  theme(legend.text.align = 0) + 
  xlab("PC1 ()") +
  ylab("PC2 (1.9%") + 
  labs(x = paste0("PC1 (", round(variance_explained[1], 3)*100, "%)"), y = paste0("PC2 (", round(variance_explained[2], 2)*100, "%)")) +
   annotate("text", x = 0.18, y = 0.28, label = paste("# sites =", num_sites), hjust = 1.1, vjust = -1.5, size = 3, color = "black") +
  geom_text_repel(aes(label=ID), size =1.5, nudge_y = 0.03, max.overlaps = 30)

ggsave("ANGSD/PCAngsd_figs/AHV_PCA_240121.png", device = "png", width = 15, height = 7, units = "cm")

#  geom_text_repel(aes(label=ID), size =1.5, nudge_y = 0.03, max.overlaps = 30)
#ggtitle("AVH1.selected.sites.co0.15x") +
#geom_text_repel(aes(label=ID), size =3, nudge_y = 0.03)

## older
AVHF1pop <- read.table("ANGSD/angsd_out/AHVF1_231220.pop.info")
cov <- as.matrix(read.table("ANGSD/angsd_out/PCAngsd/AHVF1_231220.PCAngsd.cov"))
e <- eigen(cov)

PC1 <- e$vectors[,1]
PC2 <- e$vectors[,2]
spp <- AVHF1pop[,2]
ID <- AVHF1pop[,1]
PC12.AVHF1.co0.15 <- data.frame(PC1, PC2, ID, spp)

# Calculate the proportion of variance explained by each PC
variance_explained <- e$values / sum(e$values)
variance_explained[2] # only like 2%, so very little and can ignore

ggplot(PC12.AVHF1.co0.15, aes(x=PC1, y=PC2, fill=spp)) +
  geom_point(size = 3, pch = 21, colour="black") +
  scale_fill_manual(values=c("#CC79A7", "#93c572","#56B4E9", "#FFD700"), name=paste("Species"), labels=c(expression(italic("C. allenii")), "F1","hybrid", expression(italic("C. villosissimus")))) +
  scale_x_reverse() +
  theme_bw() +
  theme(legend.text.align = 0) + 
  geom_text_repel(aes(label=ID), size =1.5, nudge_y = 0.03, max.overlaps = 30)
#ggtitle("AVH1.selected.sites.co0.15x") +
#geom_text_repel(aes(label=ID), size =3, nudge_y = 0.03)
 
  
## example of how to add repelled text labels:
#library(ggrepel)
#ggplot(mtcars, aes(wt, mpg, label = rownames(mtcars))) +
#  geom_text_repel() +
#  geom_point(color = 'red') +
#  theme_classic(base_size = 16)
```

#### PCAs of alle, vill, hybrids, guan, and laev (ANGSD_1; only clean bams with 0.15x coverage cutoff - less than 35bp and lower than 20 mapping Q removed)

##FIX## #PCAngsd run on beagle file from angsd1 (samtools; see 'ANGSD run tracking - full dataset' in \#`[g-drive]/GitHub/seq_analysis/ANGSD/angsd_runs` with estimation of independent allele frequencies. #Plot created from output of the following code:\
\#`pcangsd.py -beagle $IN_DIR/full_AVHLG_genolike.beagle.gz -o $OUT_DIR/AVHLG_1_PCAngsd -threads 44`

```{r full_angsd1_AVHLG_PCA, echo=FALSE}
LGVHApop <- read.table("ANGSD/angsd_out/LGVHA.clean.pop.info")
cov <- as.matrix(read.table("ANGSD/angsd_out/PCAngsd/sites_0.15co_nosml_q20_LGVHA_PCAngsd.cov"))
e <- eigen(cov)

PC1 <- e$vectors[,1]
PC2 <- e$vectors[,2]
spp <- LGVHApop[,2]
ID <- LGVHApop[,1]
PC12.LGVHA.co0.15X <- data.frame(PC1, PC2, ID, spp)

library(ggplot2)
ggplot(PC12.LGVHA.co0.15X, aes(x=PC1, y=PC2)) +
  geom_point(aes(color=factor(spp))) +
  ggtitle("LGVHA.co0.15x") +
  geom_text_repel(aes(label=ID), size =3, nudge_y = 0.03) 
 
  

#labeling code to check out:
#ggplot(tab, aes(x = EV2, y = EV1, col = pop)) +
#  geom_point(alpha = 0.5, size = 5) +
#  # geom_text(aes(label=sample.id), size =3, hjust = 0, nudge_x = 0.02) +
#  # geom_text(aes(label=sample.id), size =3, position=position_jitter(width=0.2,height=0,2)) +
#  geom_text_repel(aes(label=sample.id), size =3, nudge_y = 0.03) +
#  theme_bw()
```
## Local ancestry
Generated with aHMM from ancestryinfer

```{r}
## build example aHMM output plots
# chromosome 1
chr1_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr1/19.560_HYB.posterior")
chr1_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr1/19.615_HYB.posterior")
chr1_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr1/19.727_HYB.posterior")

names(chr1_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr1_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr1_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr1_19.560_0.9co <- chr1_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr1_19.615_0.9co <- chr1_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr1_19.727_0.9co <- chr1_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr1_19.560 <- ggplot(chr1_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr1 26746 sites") + geom_point() + ylim(-0.01,1.01)
chr1_19.615 <- ggplot(chr1_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr1 26746 sites") + geom_point() + ylim(-0.01,1.01)
chr1_19.727 <- ggplot(chr1_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr1 26746 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 2
chr2_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr2/19.560_HYB.posterior")
chr2_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr2/19.615_HYB.posterior")
chr2_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr2/19.727_HYB.posterior")

names(chr2_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr2_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr2_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr2_19.560_0.9co <- chr2_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr2_19.615_0.9co <- chr2_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr2_19.727_0.9co <- chr2_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr2_19.560 <- ggplot(chr2_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr2 30001 sites") + geom_point() + ylim(-0.01,1.01)
chr2_19.615 <- ggplot(chr2_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr2 30001 sites") + geom_point() + ylim(-0.01,1.01)
chr2_19.727 <- ggplot(chr2_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr2 30001 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 3
chr3_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr3/19.560_HYB.posterior")
chr3_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr3/19.615_HYB.posterior")
chr3_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr3/19.727_HYB.posterior")

names(chr3_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr3_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr3_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr3_19.560_0.9co <- chr3_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr3_19.615_0.9co <- chr3_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr3_19.727_0.9co <- chr3_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr3_19.560 <- ggplot(chr3_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr3 20397 sites") + geom_point() + ylim(-0.01,1.01)
chr3_19.615 <- ggplot(chr3_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr3 20397 sites") + geom_point() + ylim(-0.01,1.01)
chr3_19.727 <- ggplot(chr3_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr3 20397 sites") + geom_point() + ylim(-0.01,1.01)

# cromosome 4
chr4_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr4/19.560_HYB.posterior")
chr4_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr4/19.615_HYB.posterior")
chr4_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr4/19.727_HYB.posterior")

names(chr4_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr4_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr4_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr4_19.560_0.9co <- chr4_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr4_19.615_0.9co <- chr4_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr4_19.727_0.9co <- chr4_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr4_19.560 <- ggplot(chr4_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr4 17932 sites") + geom_point() + ylim(-0.01,1.01)
chr4_19.615 <- ggplot(chr4_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr4 17932 sites") + geom_point() + ylim(-0.01,1.01)
chr4_19.727 <- ggplot(chr4_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr4 17932 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 5
chr5_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr5/19.560_HYB.posterior")
chr5_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr5/19.615_HYB.posterior")
chr5_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr5/19.727_HYB.posterior")

names(chr5_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr5_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr5_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr5_19.560_0.9co <- chr5_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr5_19.615_0.9co <- chr5_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr5_19.727_0.9co <- chr5_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr5_19.560 <- ggplot(chr5_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr5 24607 sites") + geom_point() + ylim(-0.01,1.01)
chr5_19.615 <- ggplot(chr5_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr5 24607 sites") + geom_point() + ylim(-0.01,1.01)
chr5_19.727 <- ggplot(chr5_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr5 24607 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 6
chr6_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr6/19.560_HYB.posterior")
chr6_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr6/19.615_HYB.posterior")
chr6_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr6/19.727_HYB.posterior")

names(chr6_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr6_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr6_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr6_19.560_0.9co <- chr6_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr6_19.615_0.9co <- chr6_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr6_19.727_0.9co <- chr6_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr6_19.560 <- ggplot(chr6_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr6 19823 sites") + geom_point() + ylim(-0.01,1.01)
chr6_19.615 <- ggplot(chr6_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr6 19823 sites") + geom_point() + ylim(-0.01,1.01)
chr6_19.727 <- ggplot(chr6_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr6 19823 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 7
chr7_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr7/19.560_HYB.posterior")
chr7_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr7/19.615_HYB.posterior")
chr7_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr7/19.727_HYB.posterior")

names(chr7_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr7_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr7_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr7_19.560_0.9co <- chr7_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr7_19.615_0.9co <- chr7_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr7_19.727_0.9co <- chr7_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr7_19.560 <- ggplot(chr7_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr7 17428 sites") + geom_point() + ylim(-0.01,1.01)
chr7_19.615 <- ggplot(chr7_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr7 17428 sites") + geom_point() + ylim(-0.01,1.01)
chr7_19.727 <- ggplot(chr7_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr7 17428 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 8
chr8_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr8/19.560_HYB.posterior")
chr8_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr8/19.615_HYB.posterior")
chr8_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr8/19.727_HYB.posterior")

names(chr8_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr8_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr8_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr8_19.560_0.9co <- chr8_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr8_19.615_0.9co <- chr8_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr8_19.727_0.9co <- chr8_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr8_19.560 <- ggplot(chr8_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr8 17570 sites") + geom_point() + ylim(-0.01,1.01)
chr8_19.615 <- ggplot(chr8_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr8 17570 sites") + geom_point() + ylim(-0.01,1.01)
chr8_19.727 <- ggplot(chr8_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr8 17570 sites") + geom_point() + ylim(-0.01,1.01)

# chromosome 9
chr9_19.560 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr9/19.560_HYB.posterior")
chr9_19.615 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr9/19.615_HYB.posterior")
chr9_19.727 <- read_tsv("aHMM/final_params_0.02er_0.0001pulse/chr9/19.727_HYB.posterior")

names(chr9_19.560) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr9_19.615) <- c("chrom", "position", "ref_homo", "het", "alt_homo")
names(chr9_19.727) <- c("chrom", "position", "ref_homo", "het", "alt_homo")

chr9_19.560_0.9co <- chr9_19.560 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr9_19.615_0.9co <- chr9_19.615 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)
chr9_19.727_0.9co <- chr9_19.727 %>% filter(ref_homo > 0.9 | alt_homo > 0.9 | het > 0.9) %>% mutate (ancestry = ref_homo+het*0.5)

chr9_19.560 <- ggplot(chr9_19.560_0.9co, aes(position, ancestry)) + ggtitle("H_19.560 chr9 17012 sites") + geom_point() + ylim(-0.01,1.01)
chr9_19.615 <- ggplot(chr9_19.615_0.9co, aes(position, ancestry)) + ggtitle("H_19.615 chr9 17012 sites") + geom_point() + ylim(-0.01,1.01)
chr9_19.727 <- ggplot(chr9_19.727_0.9co, aes(position, ancestry)) + ggtitle("H_19.727 chr9 17012 sites") + geom_point() + ylim(-0.01,1.01)

three_ex_multipanel_wild <- chr1_19.615 / chr2_19.615 / chr3_19.615 / chr4_19.615 / chr5_19.615 / chr6_19.615 / chr7_19.615 / chr8_19.615 / chr9_19.615 |  chr1_19.727 / chr2_19.727 / chr3_19.727 / chr4_19.727 / chr5_19.727 / chr6_19.727 / chr7_19.727 / chr8_19.727 / chr9_19.727 | chr1_19.560 / chr2_19.560 / chr3_19.560 / chr4_19.560 / chr5_19.560 / chr6_19.560 / chr7_19.560 / chr8_19.560 / chr9_19.560 

ggsave("aHMM/final_params_0.02er_0.0001pulse/three_ex_multipanel_wild_aHMM_plot.pdf", three_ex_multipanel_wild, units = "mm", width = 500, height = 700)

```

## ABBA/BABA

```{r}
library('pracma')
```

### Whole genome

Code to run abba baba test and calculate D-statistic with angsd:

```{bash}
# code to run abba/baba test
angsd -doAbbababa2 1 -bam BAV_clean_ordered_0.3co_240219.bam.filelist -sizeFile sizeFile.size -doCounts 1 -out lasiOUT_full_240219 -anc /hb/groups/kay_lab/genomes/C.lasius_genome_NCBI.fasta -minQ 30 -minMapQ 30 -enhance 1 -only_proper_pairs 1 -baq 1 -ref /hb/groups/kay_lab/genomes/C.lasius_genome_NCBI.fasta -blockSize 5000000 -nThreads 96

# code to calculate D-statistic and associated stats
cd ANGSD/angsd_out/doAbbababa2
Rscript estAvgError.R angsdFile="lasiOUT_full_240215" out="lasiOUT_full_240215" sizeFile=sizeFile.size nameFile=popNames.name

# --- Table of Results --- Combination  1  out of  3  ---
# --- H1= bracteatus  H2= allenii  H3= villosissimus  H4= lasius  ---
# ---------------------------------------------------------------------------------
#   Mode          |Dstat          |sd(Dstat)      |Djack          |Zscore |Pvalue
# ---------------------------------------------------------------------------------
# Observed        |1.141e-01      |1.194e-02      |1.141e-01      |9.557  |0.0e+00
# ---------------------------------------------------------------------------------
# No Trans        |1.665e-01      |1.533e-02      |1.665e-01      |10.861 |0.0e+00
# ---------------------------------------------------------------------------------
```

### Per Chromosome

Code to run abba baba test and calculate D-statistic with angsd:

```{bash}
# use array to run eparately on each of 9 chromosomes:
angsd -doAbbababa2 1 -bam BAV_clean_ordered_0.3co_240219.bam.filelist -sizeFile sizeFile.size -doCounts 1 -out lasiOUT_chr${SLURM_ARRAY_TASK_ID}_240219 -anc /hb/groups/kay_lab/genomes/C.lasius_genome_NCBI.fasta -minQ 30 -minMapQ 30 -only_proper_pairs 1 -baq 1 -ref /hb/groups/kay_lab/genomes/C.lasius_genome_NCBI.fasta -blockSize 5000000 -nThreads 96 -r Chrom${SLURM_ARRAY_TASK_ID}

# use loop to calculate D-stats with R script for each:
for i in {1..9}; do
        Rscript estAvgError.R angsdFile="lasiOUT_chr${i}_240219" out="lasiOUT_chr${i}_240219" sizeFile=sizeFile.size nameFile=popNames.name
done
```

```{r}
# Per chromosome  values and p-val correction
raw_pvals <- c(0.000079, 0.027063, 0.000168, 0.015394, 0.040503, 0.000393, 0.04996, 0.000095, 0.152879)
FDR_pvals <- p.adjust(raw_pvals, method = "BH") # I notice that the highest p-val doesn't change at all... 
FDR_pvals # 0.00042750 0.04059450 0.00050400 0.02770920 0.05207529 0.00088425 0.05620500 0.00042750 0.15287900
```

## Dsuite

### Per Chromosome Dstats and F4-ratio

```{r}
# Per chromosome values and p-val correction (OLD)
raw_pvals <- c(8.12e-07, 2.66e-06, 4.52e-13,  1.39e-10, 7.16e-09, 1.27e-07, 1.20e-06, 1.05e-10, 6.95e-05)
FDR_pvals <- p.adjust(raw_pvals, method = "BH")
```

### Plotting f_dM and d_f

#### 50 variant windows, 25 variant steps
```{r}
# Read all chromosome files into single dataframe
# Create an empty dataframe to store the combined data
LASIout_w50s25_all9 <- data.frame()

# List all the .tsv files in the directory
file_list <- list.files(path = "/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/Dsuite/Fstats_w50s25_Mappability.Depth.Transl.Filtered.LASIref_251120", pattern = "*.txt", full.names = TRUE)

# Loop through each file and read it into a temporary dataframe
for (file in file_list) {
  temp_df <- read.delim(file, sep = "\t")
  
  # Combine the temporary dataframe with the existing data
  LASIout_w50s25_all9 <- rbind(LASIout_w50s25_all9, temp_df)
}

# Plotting 50,25 f_dM
# check for failed rows
# LASIout_w50s25_all9[is.na(LASIout_w50s25_all9$f_dM), ]
# None! yay!
# remove failed rows
# LASIout_w50s25_all9 <- LASIout_w50s25_all9[!is.na(LASIout_w50s25_all9$f_dM), ]

# ### plotting f_dM across whole genome - pos and neg values ###
# # Create cutofLASIout_w50s25_all9 with increasing stringency based on percentiles of the distribution of negative values (as suggested by Duite creator, Milan Malinsky) 
# #cutoff99thPercentile <- quantile(abs(LASIout_w50s25_all9$f_dM[which(LASIout_w50s25_all9$f_dM < 0)]),0.99)
# cutoff99.9thPercentile <- quantile(abs(LASIout_w50s25_all9$f_dM[which(LASIout_w50s25_all9$f_dM < 0)]),0.999)
# #cutoff99.99thPercentile <- quantile(abs(LASIout_w50s25_all9$f_dM[which(LASIout_w50s25_all9$f_dM < 0)]),0.9999)
# 
# # create single bp location for x-axis that adds previous chromosmes lengths
# data_cum <- LASIout_w50s25_all9 %>% 
#   group_by(chr) %>% 
#   summarise(max_windowStart = max(windowStart)) %>% 
#   mutate(windowStart_add = lag(cumsum(max_windowStart), default = 0)) %>% 
#   select(chr, windowStart_add)
# 
# LASIout_w50s25_all9 <- LASIout_w50s25_all9 %>% 
#   inner_join(data_cum, by = "chr") %>% 
#   mutate(windowStart_cum = windowStart + windowStart_add)
# 
# axis_set_dsuite <- LASIout_w50s25_all9 %>% 
#   group_by(chr) %>% 
#   summarize(center = mean(windowStart_cum))
# 
# LASI_99.9thper_w50s25_f_dM <- ggplot(data=LASIout_w50s25_all9, aes(x=windowStart_cum, y=f_dM, color = as_factor(chr))) +
#   # geom_vline(xintercept = 71772467, colour = "green", linewidth =2) + # line for % chloro reabsorp
#   # geom_vline(xintercept = 83412886, colour = "purple", linewidth =2) + # line for % N reabsorption
#   # geom_vline(xintercept = c(171402076,700244502), colour = "cyan3", linewidth =2) + # lines for seed dormancy
#   # geom_vline(xintercept = 322334915, colour = "orange", linewidth =2) + # line for toughness
#   # geom_vline(xintercept = 321497235, colour = "blue", linewidth =2) +# line for growth rate
#   geom_point() +
#   coord_cartesian(xlim = c(24398, 1120430367)) +
#   geom_hline(yintercept = c(cutoff99.9thPercentile, -cutoff99.9thPercentile), linetype = "dashed", color = "red") +
#   scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_dsuite$chr)))) +
#   guides(color = "none") +
#   geom_point(data = subset(LASIout_w50s25_all9, f_dM > cutoff99.9thPercentile | f_dM < -cutoff99.9thPercentile), color = "red") +
#   scale_x_continuous(label = axis_set_dsuite$chr, breaks = axis_set_dsuite$center) +
#   labs(x = NULL, 
#        y = "f_dM") +
#   theme(axis.text=element_text(size=40),
#         axis.title=element_text(size=40),
#         plot.margin=unit(c(2,1,1,1), 'cm'))

## Plot only positive values (representing vill alle introgression)
### plotting f_dM across whole genome (positive values only) ###

# Cutoff based on 99.9th percentile of absolute negative values
cutoff99.9thPercentile <- quantile(abs(LASIout_w50s25_all9$f_dM[LASIout_w50s25_all9$f_dM < 0]), 0.999)

# Create cumulative bp position for x-axis
data_cum <- LASIout_w50s25_all9 %>%
  group_by(chr) %>%
  summarise(max_windowStart = max(windowStart)) %>%
  mutate(windowStart_add = lag(cumsum(max_windowStart), default = 0)) %>%
  select(chr, windowStart_add)

LASIout_w50s25_all9 <- LASIout_w50s25_all9 %>%
  inner_join(data_cum, by = "chr") %>%
  mutate(windowStart_cum = windowStart + windowStart_add,
         chr_label = gsub("Chrom", "", chr))

axis_set_dsuite <- LASIout_w50s25_all9 %>%
  group_by(chr) %>%
  summarize(center = mean(windowStart_cum),
            chr_label = first(chr_label))

# Filter to positive values only
LASIout_positive <- LASIout_w50s25_all9 %>% filter(f_dM > 0)

LASI_99.9thper_w50s25_f_dM <- ggplot(data = LASIout_positive, aes(x = windowStart_cum, y = f_dM, color = as_factor(chr))) +
  geom_point(size = 3) +
  coord_cartesian(xlim = c(24398, 1120430367)) +
  geom_hline(yintercept = cutoff99.9thPercentile, linetype = "dashed", color = "red") +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_dsuite$chr)))) +
  guides(color = "none") +
  geom_point(data = subset(LASIout_positive, f_dM > cutoff99.9thPercentile), color = "red") +
  scale_x_continuous(label = axis_set_dsuite$chr_label, breaks = axis_set_dsuite$center) +
  labs(x = NULL, y = "f_dM") +
  theme(axis.text = element_text(size = 30),
        axis.title = element_text(size = 30),
        plot.margin = unit(c(2, 1, 1, 1), 'cm'))

ggsave("./Dsuite/LASI_99.9thper_w50s25_f_dM_pos_only_all_chroms.png",
       plot = LASI_99.9thper_w50s25_f_dM,
       width = 24,
       height = 5.5,
       units = "in",
       dpi = 300)

### plot subset for poster
LASIout_w50s25_all9_chrs1268 <- LASIout_w50s25_all9%>% 
  filter(chr %in% c("Chrom1","Chrom2","Chrom6","Chrom8"))

cutoff99.9thPercentile_chrs1268 <- quantile(abs(LASIout_w50s25_all9_chrs1268$f_dM[which(LASIout_w50s25_all9_chrs1268$f_dM < 0)]),0.999)

LASI_99.9thper_w50s25_f_dM_chrs1268 <- ggplot(LASIout_w50s25_all9_chrs1268, aes(x=windowStart_cum, y=f_dM, color = as_factor(chr))) +
  geom_point() +
  coord_cartesian(xlim = c(24398, 1120430367)) +
  geom_hline(yintercept = c(cutoff99.9thPercentile_chrs1268, -cutoff99.9thPercentile_chrs1268), linetype = "dashed", color = "red") +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_dsuite$chr)))) +
  guides(color = "none") +
  geom_point(data = subset(LASIout_w50s25_all9_chrs1268, f_dM > cutoff99.9thPercentile_chrs1268 | f_dM < -cutoff99.9thPercentile_chrs1268), color = "red") +
  scale_x_continuous(label = axis_set_dsuite$chr, breaks = axis_set_dsuite$center) +
  labs(x = NULL, 
       y = "f_dM") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40),
        plot.margin=unit(c(2,1,1,1), 'cm'))

# # Save plot
# ggsave("Dsuite/LASIout_w50s25_f_dM_allChrs_cutoff99.9thPer.png", LASI_99.9thper_w50s25_f_dM, device = "png", width = 110, height = 7, units = "cm")

### plotting d_f ###
# Create cutofLASIout_w50s25_all9 with increasing stringency based on percentiles of the distribution of negative values (as suggested by Duite creator, Milan Malinsky) 
cutoff99thPercentile_d_f <- quantile(abs(LASIout_w50s25_all9$d_f[which(LASIout_w50s25_all9$d_f < 0)]),0.99)
cutoff99.9thPercentile_d_f <- quantile(abs(LASIout_w50s25_all9$d_f[which(LASIout_w50s25_all9$d_f < 0)]),0.999)
cutoff99.99thPercentile_d_f <- quantile(abs(LASIout_w50s25_all9$d_f[which(LASIout_w50s25_all9$d_f < 0)]),0.9999)

LASI_99.9thper_w50s25_d_f <- ggplot(data=LASIout_w50s25_all9, aes(x=windowStart_cum, y=d_f, color = as_factor(chr))) +
  # geom_vline(xintercept = 71772467, colour = "green", linewidth =2) + # line for % chloro reabsorp
  # geom_vline(xintercept = 83412886, colour = "purple", linewidth =2) + # line for % N reabsorption
  # geom_vline(xintercept = c(171402076,700244502), colour = "cyan3", linewidth =2) + # lines for seed dormancy
  # geom_vline(xintercept = 322334915, colour = "orange", linewidth =2) + # line for toughness
  # geom_vline(xintercept = 321497235, colour = "blue", linewidth =2) +# line for growth rate
  geom_point() +
  coord_cartesian(xlim = c(24398, 1120430367)) +
  geom_hline(yintercept = c(cutoff99.9thPercentile_d_f, -cutoff99.9thPercentile_d_f), linetype = "dashed", color = "red") +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_dsuite$chr)))) +
  guides(color = "none") +
  geom_point(data = subset(LASIout_w50s25_all9, d_f > cutoff99.9thPercentile_d_f | d_f < -cutoff99.9thPercentile_d_f), color = "red") +
  scale_x_continuous(label = axis_set_dsuite$chr, breaks = axis_set_dsuite$center) +
  labs(x = NULL, 
       y = "d_f") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))

### plot subset for poster
cutoff99.9thPercentile_chrs1268 <- quantile(abs(LASIout_w50s25_all9_chrs1268$d_f[which(LASIout_w50s25_all9_chrs1268$d_f < 0)]),0.999)

LASI_99.9thper_w50s25_d_f_chrs1268 <- ggplot(LASIout_w50s25_all9_chrs1268, aes(x=windowStart_cum, y=d_f, color = as_factor(chr))) +
  geom_point() +
  coord_cartesian(xlim = c(24398, 1120430367)) +
  geom_hline(yintercept = c(cutoff99.9thPercentile_chrs1268, -cutoff99.9thPercentile_chrs1268), linetype = "dashed", color = "red") +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_dsuite$chr)))) +
  guides(color = "none") +
  geom_point(data = subset(LASIout_w50s25_all9_chrs1268, d_f > cutoff99.9thPercentile_chrs1268 | d_f < -cutoff99.9thPercentile_chrs1268), color = "red") +
  scale_x_continuous(label = axis_set_dsuite$chr, breaks = axis_set_dsuite$center) +
  labs(x = NULL, 
       y = "d_f") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40),
        plot.margin=unit(c(2,1,1,1), 'cm'))

# get first and last row per chromosome:
first_last_chr <- LASIout_w50s25_all9 %>% 
  group_by(chr) %>%
  arrange(windowStart_cum) %>%
  filter(row_number()==1 | row_number()==n())

### figure out value for other chromosmes by filtering for just chrom two and returning full line or just windowStart_cum for row with closest value... where windowStart is less than X and windowEnd is greater than X, print windowStart_cum
#OR - easier: print last row of each chromosome and just add the windowStart_cum value to X

# # Save plot
# ggsave("Dsuite/LASIout_w50s25_d_f_allChrs_cutoff99.9thPer.png", LASI_99.9thper_w50s25_d_f, device = "png", width = 110, height = 7, units = "cm")

# save as 2 panel plot:
LASI_out_f_dM.d_f_2panel <- LASI_99.9thper_w50s25_f_dM/LASI_99.9thper_w50s25_d_f
ggsave("/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/Dsuite/LASIout_w50s25_f_dM.d_f_2panel_allChrs_cutoff99.9thPer_251120.png", LASI_out_f_dM.d_f_2panel, device = "png", width = 100, height = 16, units = "cm")
```
## PIXY
```{r}
# function from the pixy website (https://pixy.readthedocs.io/en/latest/plotting.html)
pixy_to_long <- function(pixy_files){
  
  pixy_df <- list()
  
  for(i in 1:length(pixy_files)){
    
    stat_file_type <- gsub(".*_|.txt", "", pixy_files[i])
    
    if(stat_file_type == "pi"){
      
      df <- read_delim(pixy_files[i], delim = "\t")
      df <- df %>%
        gather(-pop, -window_pos_1, -window_pos_2, -chromosome,
               key = "statistic", value = "value") %>%
        rename(pop1 = pop) %>%
        mutate(pop2 = NA)
      
      pixy_df[[i]] <- df
      
      
    } else{
      
      df <- read_delim(pixy_files[i], delim = "\t")
      df <- df %>%
        gather(-pop1, -pop2, -window_pos_1, -window_pos_2, -chromosome,
               key = "statistic", value = "value")
      pixy_df[[i]] <- df
      
    }
    
  }
  
  bind_rows(pixy_df) %>%
    arrange(pop1, pop2, chromosome, window_pos_1, statistic)
  
}

# Import my data
pixy_folder <- "/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/pixy/output_251124_LASIref"
pixy_files <- list.files(pixy_folder, full.names = TRUE)
pixy_df <- pixy_to_long(pixy_files)
pixy_wide <- pivot_wider(pixy_df, names_from = statistic, values_from = value) %>% 
  mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000)

# Get a table of per chromosome average Dxy, alle/vill pi, Fst, and overall averages
pixy_wide_summary <- pixy_wide %>%
  group_by(chromosome) %>%
  summarise(
    mean_avg_dxy = mean(avg_dxy, na.rm = TRUE),
    mean_avg_wc_fst = mean(avg_wc_fst, na.rm = TRUE),
    mean_avg_pi_alle = mean(avg_pi[pop1 == "alle" & is.na(pop2)], na.rm = TRUE),
    mean_avg_pi_vill = mean(avg_pi[pop1 == "vill" & is.na(pop2)], na.rm = TRUE)
  ) %>%
  bind_rows(
    pixy_wide %>%
      summarise(
        chromosome = "TOTAL",
        mean_avg_dxy = mean(avg_dxy, na.rm = TRUE),
        mean_avg_wc_fst = mean(avg_wc_fst, na.rm = TRUE),
        mean_avg_pi_alle = mean(avg_pi[pop1 == "alle" & is.na(pop2)], na.rm = TRUE),
        mean_avg_pi_vill = mean(avg_pi[pop1 == "vill" & is.na(pop2)], na.rm = TRUE)
      )
  )
```
### Pixy plotting code
```{r}
# plot with pixy's code, panel for each chromosome

pixy_labeller <- as_labeller(c(avg_pi = "pi",
                             avg_dxy = "D[XY]",
                             avg_wc_fst = "F[ST]"),
                             default = label_parsed)

# plotting summary statistics along a single chromosome
pixy_chr1 <- pixy_df %>%
  filter(chromosome == "Chrom1") %>%
  filter(statistic %in% c("avg_pi", "avg_dxy", "avg_wc_fst")) %>%
  mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000) %>%
  ggplot(aes(x = chr_position, y = value, color = statistic))+
  geom_line(linewidth = 0.25)+
  facet_grid(statistic ~ .,
             scales = "free_y", switch = "x", space = "free_x",
             labeller = labeller(statistic = pixy_labeller))+
  xlab("Position on Chromosome (Mb)")+
  ylab("Statistic Value")+
  theme_bw()+
  theme(panel.spacing = unit(0.1, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none")+
  scale_x_continuous(expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0))+
  scale_color_brewer(palette = "Set1")

pixy_chr1

# Define a list to store the ggplot objects for each chromosome
ggplot_list <- list()

# Loop through 9 chromosomes - tried a bunch working with AI to set the y-axes for all plots of the same stat to the same range but totally failed both with chatgpt and google ai
for (i in 1:9) {
  chr_name <- paste0("pixy_chr", i)
  chr_assignment <- paste0("Chrom", i)

  # Filter the data for the current chromosome
  filtered_df <- pixy_df %>%
    filter(chromosome == chr_assignment) %>%
    filter(statistic %in% c("avg_pi", "avg_dxy", "avg_wc_fst")) %>%
    mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000)
  
  # Create the ggplot for the current chromosome
  gg <- ggplot(filtered_df, aes(x = chr_position, y = value, color = statistic)) +
    geom_line(linewidth = 0.25) +
    facet_grid(statistic ~ ., scales = "free_y", switch = "x", space = "free_x",
               labeller = labeller(statistic = pixy_labeller, value = label_value)) +
    xlab(paste0("Position on Chromosome ", i, " (Mb)")) +
    ylab("Statistic Value") +
    theme_bw() +
    theme(panel.spacing = unit(0.1, "cm"), strip.background = element_blank(),
          strip.placement = "outside", legend.position = "none") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_color_brewer(palette = "Set1")

  # Add the ggplot to the list
  ggplot_list[[chr_name]] <- gg
}

ggplot_list[["pixy_chr1"]]

pdf("pixy_style_plots_all_per_chrom_20per_each_250219.pdf", height = 12, width = 12)
plot_grid(ggplot_list[["pixy_chr1"]],ggplot_list[["pixy_chr2"]],ggplot_list[["pixy_chr3"]],ggplot_list[["pixy_chr4"]],ggplot_list[["pixy_chr5"]],ggplot_list[["pixy_chr6"]],ggplot_list[["pixy_chr7"]],ggplot_list[["pixy_chr8"]],ggplot_list[["pixy_chr9"]],
          nrow = 3,
          ncol = 3)
dev.off()

### NOTE - code below might work to set y-axes consistently, but uses too much memory - could maybe try running on HPC... 
# # Define a list to store the ggplot objects for each chromosome
# ggplot_list <- list()
# 
# # Loop through 9 chromosomes
# for (i in 1:9) {
#   chr_name <- paste0("pixy_chr", i)
#   chr_assignment <- paste0("Chrom", i)
# 
#   # Filter the data for the current chromosome
#   filtered_df <- pixy_df %>%
#     filter(chromosome == chr_assignment) %>%
#     filter(statistic %in% c("avg_pi", "avg_dxy", "avg_wc_fst")) %>%
#     mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000)
# 
#   # Create the ggplot for the current chromosome
#   gg <- ggplot(filtered_df, aes(x = chr_position, y = value, color = statistic)) +
#     geom_line(linewidth = 0.25) +
#     facet_grid(statistic ~ ., scales = "free_x", switch = "x", space = "free_x",
#                labeller = labeller(statistic = pixy_labeller, value = label_value)) +
#     xlab(paste0("Position on Chromosome ", i, " (Mb)")) +
#     ylab("Statistic Value") +
#     theme_bw() +
#     theme(panel.spacing = unit(0.1, "cm"), strip.background = element_blank(),
#           strip.placement = "outside", legend.position = "none") +
#     scale_x_continuous(expand = c(0, 0)) +
#     scale_color_brewer(palette = "Set1")
# 
# 
#     # Correctly set y-axis limits *after* faceting
#     gg <- gg +
#   facet_grid(statistic ~ ., scales = "free_x", switch = "x", space = "free_x",
#              labeller = labeller(statistic = pixy_labeller, value = label_value)) +
# 
#   scale_y_continuous(expand = c(0, 0),
#                      limits = function(limits) {
#                        scale_data <- ggplot_build(gg)$data[[1]]
#                        current_facet <- unique(scale_data$statistic)
# 
#                        if (current_facet == "avg_wc_fst") {
#                          c(0, 1)
#                        } else if (current_facet %in% c("avg_pi", "avg_dxy")) {
#                          c(0, 0.1)
#                        } else {
#                          NA_real_ # For any other stats
#                        }
#                      })
# 
#   # Add the ggplot to the list
#   ggplot_list[[chr_name]] <- gg
# }
```
### My plotting to match Dsuite plots
```{r}
### MY plotting ###
# create single bp location for x-axis that adds previous chromosome lengths
pixy_data_cum <- pixy_wide %>% 
  group_by(chromosome) %>% 
  summarise(max_windowStart = max(window_pos_1)) %>% 
  mutate(windowStart_add = lag(cumsum(max_windowStart), default = 0)) %>% 
  select(chromosome, windowStart_add)

pixy_wide <- pixy_wide %>% 
  inner_join(pixy_data_cum, by = "chromosome") %>% 
  mutate(windowStart_cum = window_pos_1 + windowStart_add)

axis_set_pixy <- pixy_wide %>% 
  group_by(chromosome) %>% 
  summarize(center = mean(windowStart_cum))

# # check congruence of chromosom positions with Dsuite
# note - not sure about this and how well it is working... windows are different bc dsuite does window size by variant number while pixy does it by bp window size, I think as long as ends of chromosomes align and it is plot by bp position it should be fine... 
# I compared the centers from axis_set data frame of each and they are relatively close, but definitely not the same... 
# One issue is dquite data doesn't start and end at the ends of the chromosome at all, not sure why... (from looking at the first_last_chr positions)
# Looking at the plots, I do not think they are aligned properly bc it looks like it is just aligning based on the ends and center of the chromosomes, which don't match up... unfortunately I also use that grouping for coloring I think... I will need to see if I can map to raw bp position within each chromosome and NOT use a cumulative val since for dsuite stuff esp the cum value will be less than the real cum if the entire chromosome were included. 
#TAKEAWAY - need to redo X axis for plotting, and likely do each chromosome separately to really see how things align. WHat I have might be an okay summary plot, but need to see each chrom carefully first. 

first_last_chr_pixy <- pixy_wide %>%
  group_by(chromosome) %>%
  arrange(windowStart_cum) %>%
  filter(row_number()==1 | row_number()==n())

## Dxy
# calc dxy cutoff
cutoff99.9thPercentile.dxy <- quantile(pixy_wide$avg_dxy,0.999, na.rm = T)
# plot dxy
Dxy_plot <- ggplot(data=pixy_wide, aes(x=windowStart_cum, y=avg_dxy, color = as_factor(chromosome))) +
  # geom_vline(xintercept = 71772467, colour = "green", linewidth =2) + # line for % chloro reabsorp
  # geom_vline(xintercept = 83412886, colour = "purple", linewidth =2) + # line for % N reabsorption
  # geom_vline(xintercept = c(171402076,700244502), colour = "cyan3", linewidth =2) + # lines for seed dormancy
  # geom_vline(xintercept = 322334915, colour = "orange", linewidth =2) + # line for toughness
  # geom_vline(xintercept = 321497235, colour = "blue", linewidth =2) +# line for growth rate
  geom_point() +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_pixy$chromosome)))) +
  guides(color = "none") +
  scale_x_continuous(label = axis_set_pixy$chromosome, breaks = axis_set_pixy$center) +
  labs(x = NULL, 
       y = "Dxy") +
  geom_hline(yintercept = cutoff99.9thPercentile.dxy, linetype = "dashed", color = "red") +
  geom_point(data = subset(pixy_wide, avg_dxy > cutoff99.9thPercentile.dxy), color = "red") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))

### subset chroms plot for poster
pixy_wide_chrs1268 <- pixy_wide %>% filter(chromosome %in% c("Chrom1", "Chrom2","Chrom6", "Chrom8"))

axis_set_pixy_chrs1268 <- axis_set_pixy %>% filter(chromosome %in% c("Chrom1", "Chrom2","Chrom6", "Chrom8"))

# calc dxy cutoff
cutoff99.9thPercentile.dxy <- quantile(pixy_wide_chrs1268$avg_dxy,0.999, na.rm = T)

# plot dxy
Dxy_plot_chrs1268 <- ggplot(data=pixy_wide_chrs1268, aes(x=windowStart_cum, y=avg_dxy, color = as_factor(chromosome))) +
  geom_point() +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_pixy_chrs1268$chromosome)))) +
  guides(color = "none") +
  scale_x_continuous(label = axis_set_pixy_chrs1268$chromosome, breaks = axis_set_pixy_chrs1268$center) +
  labs(x = NULL, 
       y = "Dxy") +
  geom_hline(yintercept = cutoff99.9thPercentile.dxy, linetype = "dashed", color = "red") +
  geom_point(data = subset(pixy_wide_chrs1268, avg_dxy > cutoff99.9thPercentile.dxy), color = "red") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))


# plot pi
cutoff99.9thPercentile.pi <- quantile(pixy_wide$avg_pi,0.999, na.rm = T)
pi_plot <- ggplot(data=pixy_wide, aes(x=windowStart_cum, y=avg_pi, color = as_factor(chromosome))) +
  # geom_vline(xintercept = 71772467, colour = "green", linewidth =2) + # line for % chloro reabsorp
  # geom_vline(xintercept = 83412886, colour = "purple", linewidth =2) + # line for % N reabsorption
  # geom_vline(xintercept = c(171402076,700244502), colour = "cyan3", linewidth =2) + # lines for seed dormancy
  # geom_vline(xintercept = 322334915, colour = "orange", linewidth =2) + # line for toughness
  # geom_vline(xintercept = 321497235, colour = "blue", linewidth =2) +# line for growth rate
  geom_point() +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_pixy$chromosome)))) +
  guides(color = "none") +
  scale_x_continuous(label = axis_set_pixy$chromosome, breaks = axis_set_pixy$center) +
  labs(x = NULL, 
       y = "pi") +
  geom_hline(yintercept = cutoff99.9thPercentile.pi, linetype = "dashed", color = "red") +
  geom_point(data = subset(pixy_wide, avg_pi > cutoff99.9thPercentile.pi ), color = "red") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))

### subset chroms plot for poster
cutoff99.9thPercentile.pi_chrs1268 <- quantile(pixy_wide_chrs1268$avg_pi,0.999, na.rm = T)
pi_plot_chrs1268 <- ggplot(data=pixy_wide_chrs1268, aes(x=windowStart_cum, y=avg_pi, color = as_factor(chromosome))) +
  geom_point() +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_pixy_chrs1268$chromosome)))) +
  guides(color = "none") +
  scale_x_continuous(label = axis_set_pixy_chrs1268$chromosome, breaks = axis_set_pixy_chrs1268$center) +
  labs(x = NULL, 
       y = "pi") +
  geom_hline(yintercept = cutoff99.9thPercentile.pi_chrs1268, linetype = "dashed", color = "red") +
  geom_point(data = subset(pixy_wide_chrs1268, avg_pi > cutoff99.9thPercentile.pi_chrs1268 ), color = "red") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))

# # plot Fst
upper_99.9thPercentile.fst <- quantile(pixy_wide$avg_wc_fst,0.999, na.rm = T)
lower_99.9thPercentile.fst <- quantile(pixy_wide$avg_wc_fst, 0.001, na.rm = T)
wc_fst_plot <- ggplot(data=pixy_wide, aes(x=windowStart_cum, y=avg_wc_fst, color = as_factor(chromosome))) +
#  geom_vline(xintercept = 71772467, colour = "green", linewidth =2) + # line for % chloro reabsorp
#  geom_vline(xintercept = 83412886, colour = "purple", linewidth =2) + # line for % N reabsorption
#  geom_vline(xintercept = c(171402076,700244502), colour = "cyan3", linewidth =2) + # lines for seed dormancy
#   geom_vline(xintercept = 322334915, colour = "orange", linewidth =2) + # line for toughness
#   geom_vline(xintercept = 321497235, colour = "blue", linewidth =2) +# line for growth rate
  geom_point() +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_pixy$chromosome)))) +
  guides(color = "none") +
  scale_x_continuous(label = axis_set_pixy$chromosome, breaks = axis_set_pixy$center) +
  labs(x = NULL, y = "Fst") +
  geom_hline(yintercept = c(upper_99.9thPercentile.fst, lower_99.9thPercentile.fst), linetype = "dashed", color = "red") +
  geom_point(data = subset(pixy_wide, avg_wc_fst > upper_99.9thPercentile.fst | avg_wc_fst < lower_99.9thPercentile.fst), color = "red") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))

### subset chroms plot for poster
upper_99.9thPercentile.fst <- quantile(pixy_wide_chrs1268$avg_wc_fst,0.999, na.rm = T)
lower_99.9thPercentile.fst <- quantile(pixy_wide_chrs1268$avg_wc_fst, 0.001, na.rm = T)
wc_fst_plot_chrs1268 <- ggplot(data=pixy_wide_chrs1268, aes(x=windowStart_cum, y=avg_wc_fst, color = as_factor(chromosome))) +
  geom_point() +
  scale_color_manual(values = rep(c("black", "darkgray"), unique(length(axis_set_pixy_chrs1268$chromosome)))) +
  guides(color = "none") +
  scale_x_continuous(label = axis_set_pixy_chrs1268$chromosome, breaks = axis_set_pixy_chrs1268$center) +
  labs(x = NULL, y = "Fst") +
  geom_hline(yintercept = c(upper_99.9thPercentile.fst, lower_99.9thPercentile.fst), linetype = "dashed", color = "red") +
  geom_point(data = subset(pixy_wide_chrs1268, avg_wc_fst > upper_99.9thPercentile.fst | avg_wc_fst < lower_99.9thPercentile.fst), color = "red") +
  theme(axis.text=element_text(size=40),
        axis.title=element_text(size=40))


# 3 panel pixy only
#Dxy_plot/pi_plot/wv_fst_plot

# 4 panel Dsuite and pixy:
combo_Dsuite_pixy_4panel <- LASI_99.9thper_w50s25_f_dM/LASI_99.9thper_w50s25_d_f/Dxy_plot/pi_plot + plot_annotation(tag_levels = 'A', tag_suffix = ')') & 
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 40, hjust = 0, vjust = 0))

ggsave("combo_Dsuite_pixy_4panel_250219.png", combo_Dsuite_pixy_4panel, device = "png", width = 100, height = 36, units = "cm")

# 5 panel with fst
combo_Dsuite_pixy_5panel <- LASI_99.9thper_w50s25_f_dM/LASI_99.9thper_w50s25_d_f/Dxy_plot/pi_plot/wc_fst_plot + plot_annotation(tag_levels = 'A', tag_suffix = ')') & 
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 40, hjust = 0, vjust = 0))

ggsave("combo_Dsuite_pixy_5panel_250219.png", combo_Dsuite_pixy_5panel, device = "png", width = 120, height = 50, units = "cm")
```

```{r}
# Chromosomes 1, 2, 6, 8 subset plots
combo_Dsuite_pixy_5panel_chrs1268 <- LASI_99.9thper_w50s25_f_dM_chrs1268/LASI_99.9thper_w50s25_d_f_chrs1268/Dxy_plot_chrs1268/pi_plot_chrs1268/wc_fst_plot_chrs1268 + plot_annotation(tag_levels = 'A', tag_suffix = ')') & 
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 40, hjust = 0, vjust = 0))

ggsave("combo_Dsuite_pixy_5panel_chroms_1268_subset.png", combo_Dsuite_pixy_5panel_chrs1268, device = "png", width = 120, height = 50, units = "cm")
```

## Splines w QTL
### Dsuite
#### import data
```{r}
# Read all chromosome files into single dataframe
# Create an empty dataframe to store the combined data
LASIout_w50s25_all9 <- data.frame()

# List all the .tsv files in the directory
file_list <- list.files(path = "/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/Dsuite/Fstats_w50s25_Mappability.Depth.Transl.Filtered.LASIref_251120", pattern = "*.txt", full.names = TRUE)

# Loop through each file and read it into a temporary dataframe
for (file in file_list) {
  temp_df <- read.delim(file, sep = "\t")
  
  # Combine the temporary dataframe with the existing data
  LASIout_w50s25_all9 <- rbind(LASIout_w50s25_all9, temp_df)
}

# Plotting 50,25 f_dM
# check for failed rows
# LASIout_w50s25_all9[is.na(LASIout_w50s25_all9$f_dM), ]
# None! yay!
# remove failed rows
# LASIout_w50s25_all9 <- LASIout_w50s25_all9[!is.na(LASIout_w50s25_all9$f_dM), ]

# plotting chromosomes separately
# load package to generate fine grain spline
library(GenWin)

# make df of positive values only (representing vill alle introgression rather than brac vill introgression)
pos_f_dM_chr <- LASIout_w50s25_all9 %>%
  select("chr","windowStart","windowEnd","f_dM") %>%
  mutate(f_dM_pos = ifelse(f_dM > 0, f_dM, NA_real_))

# code directly from GenWin example:
#chr6Spline <- splineAnalyze(Y=sub6$Fst,map=sub6$Position,smoothness=100,
#plotRaw=TRUE,plotWindows=TRUE,method=4)
# pos_f_dM_chr1 <- pos_f_dM_chr %>% 
#   filter(chr == "Chrom1")

# Keep only finite values (remove NAs)
pos_f_dM_clean <- pos_f_dM_chr[is.finite(pos_f_dM_chr$f_dM_pos), ]
# make chromosome number only col:
pos_f_dM_clean <- pos_f_dM_clean %>%
  mutate(chrN = as.numeric(str_extract(chr, "\\d+")))
```
#### Compute splines
```{r}
# Split data into list of chromosomes (names are chromosome numbers)
chr_list <- split(pos_f_dM_clean, pos_f_dM_clean$chrN)

# Initialize a results list with the same chromosome names
f_dM_spline_list <- vector("list", length(chr_list))
names(f_dM_spline_list) <- names(chr_list)

# Loop over actual chromosome names instead of 1:9
for (chr in names(chr_list)) {

  chr_df <- chr_list[[chr]]
  chr_df <- chr_df[is.finite(chr_df$f_dM_pos), ]

  message("Running spline for chromosome ", chr)

  f_dM_spline_list[[chr]] <- splineAnalyze(
    Y = chr_df$f_dM_pos,
    map = chr_df$windowStart,
    smoothness = 500000,
    plotRaw = FALSE,
    plotWindows = FALSE,
    method = 4
  )

}

# across all chroms
# Create a list of windowData tables with WindowMid added
windowed_f_dM_spline_out <- lapply(f_dM_spline_list, function(spline_obj) {

  df <- spline_obj$windowData
  df$WindowMid <- (df$WindowStart + df$WindowStop) / 2
  df
})

# Name them the same way
names(windowed_f_dM_spline_out) <- names(f_dM_spline_list)
# example of accessing one table:
# window_f_dM_chr1 <- windowed_f_dM_spline_out$chr1   # dataframe for Chrom1
# str(window_f_dM_chr1)
```
#### plot
```{r}

# plotting function to add average value line over raw data
plot_chr_combined <- function(raw_df, win_df) {
  ggplot(raw_df, aes(x = windowStart)) +
    # Raw f_dM values
    geom_point(
      aes(y = pmax(f_dM, 0)),
      size = 1,
      alpha = 0.4,
      color = "grey40"
    ) +
    # MeanY line (no scaling needed)
    geom_line(
      data = win_df,
      aes(x = WindowMid, y = MeanY),
      color = "darkred",
      linewidth = 0.7
    ) +
    geom_point(
      data = win_df,
      aes(x = WindowMid, y = MeanY),
      color = "darkred",
      size = 1.5
    ) +
    labs(
      x = "Position (bp)",
      y = "f_dM"
    ) +
    theme_bw()
}

# test on one chromosome
plot_chr_combined(
  raw_df = pos_f_dM_clean[pos_f_dM_clean$chrN == 1, ],
  win_df = windowed_spline_out$chr1
)

# Make chroms list
chroms <- sort(unique(pos_f_dM_clean$chrN))

# Make a plot for each chromosome
plots <- lapply(chroms, function(i) {
  
  raw_df_i <- pos_f_dM_clean[pos_f_dM_clean$chrN == i, ]
  win_df_i <- windowed_f_dM_spline_out[[paste0(i)]]
  
  plot_chr_combined(
    raw_df = raw_df_i,
    win_df = win_df_i
  )
})

# Optionally name them
names(plots) <- paste0("chr", chroms)

# check/test
plots[[1]] 

### Add lines for QTL peaks and credible intervals 
# import QTL peak locations and CIs from generated with rqtl script
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs_bp_cm.251130.csv", header = T)

# remove the huge CI secondary Nreabsorption, SD, and GR peaks:
all_QTL_CI <- all_QTL_CI %>%
  filter(!chr==2 | !trait =="Nreabs") %>%
  filter(!chr==5 | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)
#library(colorblindcheck)
#colorblindcheck::palette_plot(c("#0072B2", "#CC79A7", "#8250C4", "#E69F00", "#009E73"))

# list of chromosome numbers
chrom_list <- sort(unique(pos_f_dM_clean$chrN))

# helper function to add QTL peaks and CIs
add_qtl_layers <- function(base_plot, qtl_rows, trait_colors) {

  if (nrow(qtl_rows) == 0) return(base_plot)   # nothing to add

  base_plot +
    # CI shading
    geom_rect(
      data = qtl_rows,
      aes(
        xmin = lower95CI_bp,
        xmax = upper95CI_bp,
        ymin = -Inf,
        ymax = Inf,
        fill = trait
      ),
      inherit.aes = FALSE,
      alpha = 0.2
    ) +
    # Vertical peak lines
    geom_vline(
      data = qtl_rows,
      aes(xintercept = peak_bp, color = trait),
      linewidth = 1,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = trait_colors) +
    scale_color_manual(values = trait_colors)
}

# build empty list to store plots
final_plots <- list()

# apply new function across all chromosomes
for (i in seq_along(chrom_list)) {
  
  chr_num <- chrom_list[i]            
  base_plot <- plots[[i]]                # spline-only plot
  
  qtl_here <- all_QTL_CI %>% 
    filter(chr == chr_num)              # get QTL rows for this chromosome
  
  plot_i <- add_qtl_layers(base_plot, qtl_here, trait_colors)

  plot_i <- plot_i + ggtitle(paste("Chromosome", chr_num))
  
  # save
  final_plots[[i]] <- plot_i
}

# name plots by chromosome name
names(final_plots) <- paste0("Chrom", chrom_list)

final_plots$Chrom5 # to call a single plot to check 

# make into a multipanel and save
f_dM_LASI_w50s25_sep_chrs_w.QTL <- final_plots$Chrom1 + final_plots$Chrom2 + final_plots$Chrom3 + final_plots$Chrom5 + final_plots$Chrom6 + final_plots$Chrom7 + final_plots$Chrom8 + final_plots$Chrom9

ggsave("Dsuite/LASIout_w50s25_f_dM_separate_chrs_w.500kb.spline_7QTL_narrowed.CIs.png", f_dM_LASI_w50s25_sep_chrs_w.QTL, device = "png", width = 50, height = 20, units = "cm")
```

### pixy
#### import data
```{r}
# function from the pixy website (https://pixy.readthedocs.io/en/latest/plotting.html)
pixy_to_long <- function(pixy_files){
  
  pixy_df <- list()
  
  for(i in 1:length(pixy_files)){
    
    stat_file_type <- gsub(".*_|.txt", "", pixy_files[i])
    
    if(stat_file_type == "pi"){
      
      df <- read_delim(pixy_files[i], delim = "\t")
      df <- df %>%
        gather(-pop, -window_pos_1, -window_pos_2, -chromosome,
               key = "statistic", value = "value") %>%
        rename(pop1 = pop) %>%
        mutate(pop2 = NA)
      
      pixy_df[[i]] <- df
      
      
    } else{
      
      df <- read_delim(pixy_files[i], delim = "\t")
      df <- df %>%
        gather(-pop1, -pop2, -window_pos_1, -window_pos_2, -chromosome,
               key = "statistic", value = "value")
      pixy_df[[i]] <- df
      
    }
    
  }
  
  bind_rows(pixy_df) %>%
    arrange(pop1, pop2, chromosome, window_pos_1, statistic)
  
}

# Import my data
pixy_folder <- "/Users/juliaharencar/Documents/Github/Population_genetics_vill_alle/pixy/output_251124_LASIref"
pixy_files <- list.files(pixy_folder, full.names = TRUE)
pixy_df <- pixy_to_long(pixy_files)
pixy_wide <- pivot_wider(pixy_df, names_from = statistic, values_from = value) %>% 
  mutate(chr_position = ((window_pos_1 + window_pos_2)/2)/1000000)
# make chromosome number only col:
pixy_wide <- pixy_wide %>%
  mutate(chrN = as.numeric(str_extract(chromosome, "\\d+")))
```
#### compute splines
```{r}
# Compute spline
# Make function to run splineAnalyze on cleaned data for each chromosome
# Split into a list of dataframes, one per chromosome
pixy_chr_list <- split(pixy_wide, pixy_wide$chrN)
# pixy_chr_list[[1]] is Chrom1

# run splineAnalyze on each chromosome:
dxy_spline_list <- vector("list", 9)

for (i in 1:9) {

  chr_df <- pixy_chr_list[[i]]
  chr_df <- chr_df[is.finite(chr_df$avg_dxy), ]

  message("Running spline for chromosome ", i)

  dxy_spline_list[[i]] <- splineAnalyze(
    Y = chr_df$avg_dxy,
    map = chr_df$window_pos_1,
    smoothness = 500000,
    plotRaw = FALSE,
    plotWindows = FALSE,
    method = 4
  )
}

names(dxy_spline_list) <- paste0("chr", 1:9)

# across all chroms
# Create a list of windowData tables with WindowMid added
windowed_Dxy_spline_out <- lapply(dxy_spline_list, function(spline_obj) {

  df <- spline_obj$windowData
  df$WindowMid <- (df$WindowStart + df$WindowStop) / 2
  df
})

# Name them the same way
names(windowed_Dxy_spline_out) <- names(dxy_spline_list)
# example of accessing one table:
window_dxy_chr1 <- windowed_Dxy_spline_out$chr1   # dataframe for Chrom1
(window_dxy_chr1)
```
#### plot
```{r}
# # function to plot spline over raw data
# plot_chr_combined <- function(raw_df, win_df) {
#   
#   # Range for scaling
#   y1_min <- 0
#   y1_max <- max(raw_df$avg_dxy, na.rm = TRUE)
#   
#   y2_min <- min(win_df$Wstat, na.rm = TRUE)
#   y2_max <- max(win_df$Wstat, na.rm = TRUE)
#   
#   # scale() converts Wstat → dxy-like scale
#   scale_Wstat <- function(x) {
#     (x - y2_min) / (y2_max - y2_min) * (y1_max - y1_min) + y1_min
#   }
#   
#   # inverse scale: dxy-like → original Wstat
#   inv_scale <- function(x) {
#     (x - y1_min) / (y1_max - y1_min) * (y2_max - y2_min) + y2_min
#   }
#   
#   ggplot(raw_df, aes(x = window_pos_1)) +
#     
#     # Raw dxy values
#     geom_point(aes(y = avg_dxy),
#                size = 1, alpha = 0.4, color = "grey40") +
#     
#     # Wstat spline (scaled)
#     geom_line(data = win_df,
#               aes(x = WindowMid, y = scale_Wstat(Wstat)),
#               color = "darkred", linewidth = 0.7) +
#     geom_point(data = win_df,
#                aes(x = WindowMid, y = scale_Wstat(Wstat)),
#                color = "darkred", size = 1.5) +
# 
#     # Labels and second axis
#     scale_y_continuous(
#       name = "Dxy",
#       sec.axis = sec_axis(~ inv_scale(.),
#                           name = "Wstat (window deviation)")
#     ) +
#     labs(x = "Position (bp)") +
#     theme_bw() +
#     theme(
#       axis.title.y.left = element_text(color = "black"),
#       axis.title.y.right = element_text(color = "darkred"),
#       axis.text.y.right  = element_text(color = "darkred")
#     )
# }

# # test on one chromosome
# plot_chr_combined(
#   raw_df = pixy_wide[pixy_wide$chrN == 1, ],
#   win_df = windowed_Dxy_spline_out$chr1
# )

# plotting function to add average value line over raw data
plot_chr_combined <- function(raw_df, win_df) {
  ggplot(raw_df, aes(x = window_pos_1)) +
    # Raw f_dM values
    geom_point(
      aes(y = pmax(avg_dxy, 0)),
      size = 1,
      alpha = 0.4,
      color = "grey40"
    ) +
    # MeanY line (no scaling needed)
    geom_line(
      data = win_df,
      aes(x = WindowMid, y = MeanY),
      color = "darkred",
      linewidth = 0.7
    ) +
    geom_point(
      data = win_df,
      aes(x = WindowMid, y = MeanY),
      color = "darkred",
      size = 1.5
    ) +
    labs(
      x = "Position (bp)",
      y = "Dxy"
    ) +
    theme_bw()
}

# test on one chromosome
plot_chr_combined(
  raw_df = pixy_wide[pixy_wide$chrN == 1, ],
  win_df = windowed_Dxy_spline_out$chr1
)

chroms <- 1:9

# Make a plot for each chromosome
plots <- lapply(chroms, function(i) {
  
  raw_df_i <- pixy_wide[pixy_wide$chrN == i, ]
  win_df_i <- windowed_Dxy_spline_out[[paste0("chr", i)]]
  
  plot_chr_combined(
    raw_df = raw_df_i,
    win_df = win_df_i
  )
})

# Optionally name them
names(plots) <- paste0("chr", chroms)

# check/test
#plots[[1]] # works!

### Add lines for QTL peaks and credible intervals 
# import QTL peak locations and CIs from generated with rqtl script
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs_bp_cm.251130.csv", header = T)

# remove the huge CI secondary Nreabsorption, SD, and GR peaks:
all_QTL_CI <- all_QTL_CI %>% 
  filter(!chr==2 | !trait =="Nreabs") %>%
  filter(!chr==5 | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)
#library(colorblindcheck)
#colorblindcheck::palette_plot(c("#0072B2", "#CC79A7", "#8250C4", "#E69F00", "#009E73"))

# list of chromosome numbers
chrom_list <- sort(unique(pixy_wide$chrN))

# helper function to add QTL peaks and CIs
add_qtl_layers <- function(base_plot, qtl_rows, trait_colors) {

  if (nrow(qtl_rows) == 0) return(base_plot)   # nothing to add

  base_plot +
    # CI shading
    geom_rect(
      data = qtl_rows,
      aes(
        xmin = lower95CI_bp,
        xmax = upper95CI_bp,
        ymin = -Inf,
        ymax = Inf,
        fill = trait
      ),
      inherit.aes = FALSE,
      alpha = 0.2
    ) +
    # Vertical peak lines
    geom_vline(
      data = qtl_rows,
      aes(xintercept = peak_bp, color = trait),
      linewidth = 1,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = trait_colors) +
    scale_color_manual(values = trait_colors)
}

# build empty list to store plots
final_plots <- list()

# apply new function across all chromosomes
for (i in seq_along(chrom_list)) {
  
  chr_num <- chrom_list[i]            
  base_plot <- plots[[i]]                # spline-only plot
  
  qtl_here <- all_QTL_CI %>% 
    filter(chr == chr_num)              # get QTL rows for this chromosome
  
  plot_i <- add_qtl_layers(base_plot, qtl_here, trait_colors)

  plot_i <- plot_i + ggtitle(paste("Chromosome", chr_num))
  
  # save
  final_plots[[i]] <- plot_i
}

# name plots by chromosome name
names(final_plots) <- paste0("Chrom", chrom_list)

final_plots$Chrom1 # to call a single plot to check 

# make into a multipanel and save
Dxy_LASI_sep_chrs_w.QTL <- final_plots$Chrom1 + final_plots$Chrom2 + final_plots$Chrom3 + final_plots$Chrom5 + final_plots$Chrom6 + final_plots$Chrom7 + final_plots$Chrom8 + final_plots$Chrom9

ggsave("pixy/Dxy_10kb_LASIout_separate_chrs_w.500kb.spline_7QTL_narrowed.CIs.png", Dxy_LASI_sep_chrs_w.QTL, device = "png", width = 50, height = 20, units = "cm")



########### same as above but for Fst rather than Dxy ################
# plotting function to add average value line over raw data
plot_chr_combined <- function(raw_df, win_df) {
  ggplot(raw_df, aes(x = window_pos_1)) +
    # Raw f_dM values
    geom_point(
      aes(y = pmax(avg_wc_fst, 0)),
      size = 1,
      alpha = 0.4,
      color = "grey40"
    ) +
    # MeanY line (no scaling needed)
    geom_line(
      data = win_df,
      aes(x = WindowMid, y = MeanY),
      color = "darkred",
      linewidth = 0.7
    ) +
    geom_point(
      data = win_df,
      aes(x = WindowMid, y = MeanY),
      color = "darkred",
      size = 1.5
    ) +
    labs(
      x = "Position (bp)",
      y = "Fst"
    ) +
    theme_bw()
}

# test on one chromosome
plot_chr_combined(
  raw_df = pixy_wide[pixy_wide$chrN == 1, ],
  win_df = windowed_Fst_spline_out$chr1
)

chroms <- 1:9

# Make a plot for each chromosome
plots <- lapply(chroms, function(i) {
  
  raw_df_i <- pixy_wide[pixy_wide$chrN == i, ]
  win_df_i <- windowed_Fst_spline_out[[paste0("chr", i)]]
  
  plot_chr_combined(
    raw_df = raw_df_i,
    win_df = win_df_i
  )
})

# Optionally name them
names(plots) <- paste0("chr", chroms)

# check/test
#plots[[1]] # works!

### Add lines for QTL peaks and credible intervals 
# import QTL peak locations and CIs from generated with rqtl script
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs_bp_cm.251130.csv", header = T)

# remove the huge CI secondary Nreabsorption, SD, and GR peaks:
all_QTL_CI <- all_QTL_CI %>% 
  filter(!chr==2 | !trait =="Nreabs") %>%
  filter(!chr==5 | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)
#library(colorblindcheck)
#colorblindcheck::palette_plot(c("#0072B2", "#CC79A7", "#8250C4", "#E69F00", "#009E73"))

# list of chromosome numbers
chrom_list <- sort(unique(pixy_wide$chrN))

# helper function to add QTL peaks and CIs
add_qtl_layers <- function(base_plot, qtl_rows, trait_colors) {

  if (nrow(qtl_rows) == 0) return(base_plot)   # nothing to add

  base_plot +
    # CI shading
    geom_rect(
      data = qtl_rows,
      aes(
        xmin = lower95CI_bp,
        xmax = upper95CI_bp,
        ymin = -Inf,
        ymax = Inf,
        fill = trait
      ),
      inherit.aes = FALSE,
      alpha = 0.2
    ) +
    # Vertical peak lines
    geom_vline(
      data = qtl_rows,
      aes(xintercept = peak_bp, color = trait),
      linewidth = 1,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = trait_colors) +
    scale_color_manual(values = trait_colors)
}

# build empty list to store plots
final_plots <- list()

# apply new function across all chromosomes
for (i in seq_along(chrom_list)) {
  
  chr_num <- chrom_list[i]            
  base_plot <- plots[[i]]                # spline-only plot
  
  qtl_here <- all_QTL_CI %>% 
    filter(chr == chr_num)              # get QTL rows for this chromosome
  
  plot_i <- add_qtl_layers(base_plot, qtl_here, trait_colors)

  plot_i <- plot_i + ggtitle(paste("Chromosome", chr_num))
  
  # save
  final_plots[[i]] <- plot_i
}

# name plots by chromosome name
names(final_plots) <- paste0("Chrom", chrom_list)

final_plots$Chrom1 # to call a single plot to check 

# make into a multipanel and save
Fst_LASI_sep_chrs_w.QTL <- final_plots$Chrom1 + final_plots$Chrom2 + final_plots$Chrom3 + final_plots$Chrom5 + final_plots$Chrom6 + final_plots$Chrom7 + final_plots$Chrom8 + final_plots$Chrom9

ggsave("pixy/Fst_10kb_LASIout_separate_chrs_w.500kb.spline_7QTL_narrowed.CIs.png", Fst_LASI_sep_chrs_w.QTL, device = "png", width = 50, height = 20, units = "cm")

```

### combined
```{r}

## NOTE - works up to making plots with the wstat for both stats per chrom, but code to add QTL with CIs doesn't work 
# test joint plot
f_dM_wstat_df <- windowed_f_dM_spline_out[["1"]] %>%
  select(WindowMid, f_dM_Wstat = Wstat) %>%
  mutate(stat = "f_dM_Wstat") %>%
  rename(value = f_dM_Wstat)

dxy_wstat_df <- windowed_Dxy_spline_out[["chr2"]] %>%
  select(WindowMid, Dxy_Wstat = Wstat) %>%   
  mutate(stat = "Dxy_Wstat") %>%
  rename(value = Dxy_Wstat)

# Combine into one long data frame
combined_df <- bind_rows(f_dM_wstat_df, dxy_wstat_df)

# Plot
ggplot(combined_df, aes(x = WindowMid, y = value, color = stat)) +
  geom_line(size = 0.8) +
  geom_point(size = 1.5, alpha = 0.7) +
  labs(
    x = "Position (bp)",
    y = "Statistic value",
    color = "Statistic") +
  theme_bw()

# add QTL

### Add lines for QTL peaks and credible intervals 
# import QTL peak locations and CIs from generated with rqtl script
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs.251122.csv", header = T)

# remove the huge CI secondary Nreabsorption, SD, and GR peaks:
all_QTL_CI <- all_QTL_CI %>% 
  filter(!chr=="Chrom2" | !trait =="Nreabs") %>% 
  filter(!chr=="Chrom4" | !trait =="SD") %>%
  filter(!chr=="Chrom5" | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)
#library(colorblindcheck)
#colorblindcheck::palette_plot(c("#0072B2", "#CC79A7", "#8250C4", "#E69F00", "#009E73"))

# list of chromosome names
chrom_list <- unique(pos_f_dM_chr$chr)
# helper function to add QTL peaks and CIs
add_qtl_layers <- function(base_plot, qtl_rows, trait_colors) {

  if (nrow(qtl_rows) == 0) return(base_plot)   # nothing to add

  base_plot +
    # CI shading
    geom_rect(
      data = qtl_rows,
      aes(
        xmin = lower95CI_bp,
        xmax = upper95CI_bp,
        ymin = -Inf,
        ymax = Inf,
        fill = trait
      ),
      inherit.aes = FALSE,
      alpha = 0.2
    ) +
    # Vertical peak lines
    geom_vline(
      data = qtl_rows,
      aes(xintercept = peak_bp, color = trait),
      linewidth = 1,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = trait_colors) +
    scale_color_manual(values = trait_colors)
}

# build empty list to store plots
final_plots <- list()

# apply new function across all chromosomes
for (i in seq_along(chrom_list)) {
  
  chr_name <- chrom_list[i]              # e.g. "Chrom6"
  base_plot <- plots[[i]]                # spline-only plot
  
  qtl_here <- all_QTL_CI %>% 
    filter(chr == chr_name)              # get QTL rows for this chromosome
  
  final_plots[[i]] <- add_qtl_layers(base_plot, qtl_here, trait_colors)
}

# name plots by chromosome name
names(final_plots) <- chrom_list

final_plots$Chrom1 # to call a single plot to check 





############# PICKUP ###########
# TEST the below to plot all with QTL:
# --- 1. Prepare chromosome list (skip missing chromosomes automatically) ---
chrom_list <- intersect(unique(pos_f_dM_chr$chr), names(plots))  # only chromosomes with plots

# --- 2. Helper function to add QTL peaks and CIs ---
add_qtl_layers <- function(base_plot, qtl_rows, trait_colors) {
  if (nrow(qtl_rows) == 0) return(base_plot)   # nothing to add

  base_plot +
    # CI shading
    geom_rect(
      data = qtl_rows,
      aes(
        xmin = lower95CI_bp,
        xmax = upper95CI_bp,
        ymin = -Inf,
        ymax = Inf,
        fill = trait
      ),
      inherit.aes = FALSE,
      alpha = 0.2
    ) +
    # Vertical peak lines
    geom_vline(
      data = qtl_rows,
      aes(xintercept = peak_bp, color = trait),
      linewidth = 1,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = trait_colors) +
    scale_color_manual(values = trait_colors)
}

# --- 3. Build plots with QTL ---
final_plots <- list()

for (chr_name in chrom_list) {

  # Base spline + statistic plot for this chromosome
  base_plot <- plots[[chr_name]]   # now we use the actual name

  # QTL for this chromosome
  qtl_here <- all_QTL_CI %>%
    filter(chr == chr_name)

  # Add QTL layers
  final_plots[[chr_name]] <- add_qtl_layers(base_plot, qtl_here, trait_colors)
}

# final_plots is now named by chromosome
names(final_plots)  # check
final_plots$Chrom2  # example plot
```

### windowed avgs
Calculate average f_dM and Dxy in cM windows and plot distribution, then plot lines with QTL window averages over that.
#### calc cM
```{r}
# first, import cM positions from lepmap; will only be for variant sites, but hoping maybe we can impute for the rest somehow...?

# make a list of the files with marker and cM distance info
lepmap_files <- list.files("/Users/juliaharencar/Documents/Github/alle_vill_QTL/lepmap3/best_iter_ordermarkers/", pattern = "*_clean_LodLim11_theta0.001_allF2s_0.5covCO_withRef.txt", full.names = TRUE)

# initiate an empty df
lepmap_df <- c()

# Use lapply to import marker data from all chromosomes into one df
lepmap_df <- do.call(rbind, lapply(lepmap_files, function(f) {
  markers <- read.table(f, sep = "\t", header = FALSE, stringsAsFactors = FALSE)
  names(markers) <- c("marker_num", "cM_pos", "chrom", "bp_pos")
  return(markers)
}))
# remove chromosome 4 because of poor genetic map and absence of QTL and remove arbitrary marker number col
bp_to_cm <- lepmap_df %>% 
  filter(chrom != 4) %>% # remove chromosome 4 becasue bad genetic map
  filter(!(chrom == 2 & bp_pos > 1.37e+08 & bp_pos < 1.4e+08)) %>% #remove 3 erroneous markers on chromosome 2
  select("chrom", "bp_pos", "cM_pos")

# clean up maps to get data that are monotonic and can be used to generate an interpolation function:
# check maps - chr4 was junk and that is why we removed it above from this analysis
# chromosome 2 has a few markers off the curve that also get removed above 
# bp_to_cm %>% filter(chrom == 2) %>% 
#    ggplot(., aes(x = bp_pos, y = cM_pos)) +
#    geom_point(alpha = 0.4, size = 1) +4
#    facet_wrap(~ chrom, scales = "free_x") +
#    labs(
#      x = "Physical position (bp)",
#      y = "Genetic position (cM)",
#      title = "Genetic map check: cM vs bp"
#    ) +
#    theme_bw()

# Chromosomes to invert
chrom_invert <- c(2,3,7,9)

bp_to_cm_fixed <- bp_to_cm %>%
  group_by(chrom) %>%
  mutate(cM_pos = ifelse(chrom %in% chrom_invert,
                         max(cM_pos) - cM_pos,
                         cM_pos)) %>%
  ungroup()

# apply isotonic regression to get smooth function line that is monotonically increasing
apply_iso <- function(df, bp_col = "bp_pos", cm_col = "cM_pos") {
  df <- df %>% arrange(.data[[bp_col]])
  y  <- df[[cm_col]]

  iso <- isoreg(y)      # monotone non-decreasing fit
  df$cM_iso <- iso$yf   # fitted (monotonic) cM
  df
}

# fit isotonic curve to each chromosome
bp_to_cm_iso <- bp_to_cm_fixed %>%
  group_by(chrom) %>%
  group_modify(~ apply_iso(.x)) %>%
  ungroup()

# plot to check
ggplot(bp_to_cm_iso, aes(x = bp_pos)) +
  geom_point(aes(y = cM_pos), size = 0.8, alpha = 0.4) +
  geom_line(aes(y = cM_iso), color = "blue", size = 0.7) +
  facet_wrap(~ chrom, scales = "free_x") +
  labs(title = "Original cM (points) vs Isotonic monotone cM (blue line)",
       x = "bp_pos", y = "cM") +
  theme_bw()

# Create a named list of functions, one per chromosome to interpolate cm from bp
bp_to_cm_funs <- bp_to_cm_iso %>%
  group_by(chrom) %>%
  group_map(~ {
    df <- .x
    # approxfun creates a function: bp -> cM_iso
    approxfun(df$bp_pos, df$cM_iso, rule = 2)  # rule=2 extends at edges
  }) 

# Name the functions by chromosome
names(bp_to_cm_funs) <- unique(bp_to_cm_iso$chrom)
# usage: E.g. bp_to_cm_funs[["2"]](1.37e8)  # returns interpolated cM
```
#### f_dM
```{r}
# remove chromosome 4 from f_dM data and translate window start bp to cm
pos_f_dM_clean <- pos_f_dM_clean %>%
  filter(chrN != 4) %>% 
  rowwise() %>%
  mutate(cM_start = bp_to_cm_funs[[as.character(chrN)]](windowStart)) %>%
  ungroup()

# bin by cM and calc f_dM averages
f_dM_summary <- pos_f_dM_clean %>%
  mutate(cM_bin = floor(cM_start)) %>%  # 1 cM bins
  group_by(chrN, cM_bin) %>%
  summarise(avg_f_dM_per_cM = mean(f_dM_pos, na.rm = TRUE),
            n_windows = n()) %>%
  ungroup()

# load QTL intervals, calc and plot average dxy in the windows
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs_bp_cm.251130.csv", header = T)

# remove the huge CI Nreabsorption, toughness, and GR peaks:
all_QTL_CI <- all_QTL_CI %>% 
  filter(!chr==2 | !trait =="Nreabs") %>% 
#  filter(!chr==6 | !trait =="SD") %>%
  filter(!chr==5 | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)

# calculate average f_dM for each peak PICKUP - fix to use cM pos not bp - bc might be cutting CI really short bc many markers at same cM position. 
qtl_f_dM_bp <- all_QTL_CI %>%
  rowwise() %>%
  mutate(
    avg_f_dM_QTL = mean(
      pos_f_dM_clean$f_dM[
        pos_f_dM_clean$chrN == chr &
        pos_f_dM_clean$windowStart <= upper95CI_bp &  # window starts before interval end
        pos_f_dM_clean$windowEnd >= lower95CI_bp    # window ends after interval start
      ],
      na.rm = TRUE
    )
  ) %>%
  ungroup()

# plot qtl peak averages on histogram of average per window values
f_dM_windows_hist_w.QTL.CIs <- ggplot() +
  # Genome-wide histogram of 1 cM avg_dxy
  geom_histogram(
    data = f_dM_summary,
    aes(x = avg_f_dM_per_cM, y = after_stat(density)),
    bins = 50,
    fill = "grey80",
    color = "black"
  ) +
  # Vertical lines at QTL bp-interval averages
  geom_vline(
    data = qtl_f_dM_bp,
    aes(xintercept = avg_f_dM_QTL, color = trait),
    linetype = "solid",
    size = 1
  ) +
  scale_color_manual(values = trait_colors) +
  labs(
    x = "average f_dM",
    y = "frequency",
    color = "Trait"
  ) +
  theme_bw()

# save plot
ggsave("f_dM_windows_hist_w.QTL.CIs.png", f_dM_windows_hist_w.QTL.CIs, device = "png", width = 20, height = 12, units = "cm")

```

#### Dxy
```{r}
# Add cM position to df with Dxy values
# first, clean/simplify full pixy df, including removal of chrom4
dxy <- pixy_wide %>% 
  filter(!is.na(avg_dxy) & chromosome != "Chrom4") %>%
  mutate(window_center = chr_position*10e5) %>% 
  select(chrN, chromosome, window_pos_1, window_pos_2, avg_dxy, window_center) %>% 
  as.data.frame()

# apply function to dxy window start points
dxy <- dxy %>%
  rowwise() %>%
  mutate(cM_window_start = bp_to_cm_funs[[as.character(chrN)]](window_pos_1)) %>%
  ungroup()

# calculate dxy in 1 cM windows
dxy_summary <- dxy %>%
  mutate(cM_bin = floor(cM_window_start)) %>%   # bin by integer cM
  group_by(chrN, cM_bin) %>%
  summarise(avg_dxy_per_cM = mean(avg_dxy, na.rm = TRUE),
            n_windows = n()) %>%
  ungroup()


# load QTL intervals, calc and plot average dxy in the windows
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs_bp_cm.251130.csv", header = T)

# remove the huge CI secondary Nreabsorption, SD, and GR peaks:
all_QTL_CI <- all_QTL_CI %>% 
  filter(!chr==2 | !trait =="Nreabs") %>% 
  filter(!chr==6 | !trait =="SD") %>%
  filter(!chr==5 | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)
 
# calculate average dxy for each peak PICKUP - fix to use cM pos not bp - bc might be cutting CI really short bc many markers at same cM position. 
qtl_dxy_bp <- all_QTL_CI %>%
  rowwise() %>%
  mutate(
    avg_dxy_QTL = mean(
      dxy$avg_dxy[
        dxy$chrN == chr &
        dxy$window_pos_1 <= upper95CI_bp &  # window starts before interval end
        dxy$window_pos_2 >= lower95CI_bp    # window ends after interval start
      ],
      na.rm = TRUE
    )
  ) %>%
  ungroup()

dxy_windows_hist_w.QTL.CIs <- ggplot() +
  # Histogram
  geom_histogram(
    data = dxy_summary,
    aes(x = avg_dxy_per_cM, y = after_stat(density)),
    bins = 50,
    fill = "grey80",
    color = "black"
  ) +
  # QTL lines (legend ON + force point glyph)
  geom_vline(
    data = qtl_dxy_bp,
    aes(xintercept = avg_dxy_QTL, color = trait),
    linetype = "solid",
    size = 1,
    show.legend = TRUE,
    key_glyph = "point"     # <-- force legend to draw using point key
  ) +
  # Colors
  scale_color_manual(values = trait_colors) +
  # Override legend aesthetics so points become squares
  guides(
    color = guide_legend(
      override.aes = list(
        shape = 15,     # square
        size = 5,
        linetype = 0    # no line in the legend
      )
    )
  ) +
  labs(
    x = "average Dxy",
    y = "frequency",
    color = "Trait"
  ) +
  theme_bw() +
  theme(
    legend.position = "right",
    legend.justification = "center"
  )


# save plot
# chr 2 SD peak is present, just under chr8 toughness peak
ggsave("dxy_windows_hist_w.QTL.CIs.png", dxy_windows_hist_w.QTL.CIs, device = "png", width = 20, height = 12, units = "cm")

### save both plots together ###
# Suppress legend for f_dM plot
p1 <- f_dM_windows_hist_w.QTL.CIs +
  guides(color = "none")

# Keep legend for dxy plot
p2 <- dxy_windows_hist_w.QTL.CIs

# Combine and collect only the legend from p2
f_dm.dxy_hist_w.QTL.CIs <-
  p1 / p2 +
  plot_layout(guides = "collect")

ggsave("dxy_windows_hist_w.6QTL.CIs.png", f_dm.dxy_hist_w.QTL.CIs, device = "png", width = 13, height = 10, units = "cm")
```
#### Fst
```{r}
# Add cM position to df with Fst values
# first, clean/simplify full pixy df, including removal of chrom4
fst <- pixy_wide %>% 
  filter(!is.na(avg_wc_fst) & chromosome != "Chrom4") %>%
  mutate(window_center = chr_position*10e5) %>% 
  select(chrN, chromosome, window_pos_1, window_pos_2, avg_wc_fst, window_center) %>% 
  as.data.frame()

# apply function to fst window start points
fst <- fst %>%
  rowwise() %>%
  mutate(cM_window_start = bp_to_cm_funs[[as.character(chrN)]](window_pos_1)) %>%
  ungroup()

# calculate fst in 1 cM windows
fst_summary <- fst %>%
  mutate(cM_bin = floor(cM_window_start)) %>%   # bin by integer cM
  group_by(chrN, cM_bin) %>%
  summarise(avg_fst_per_cM = mean(avg_wc_fst, na.rm = TRUE),
            n_windows = n()) %>%
  ungroup()


# load QTL intervals, calc and plot average fst in the windows
all_QTL_CI <- read.csv("/Users/juliaharencar/Documents/Github/alle_vill_QTL/all_QTL_CIs_bp_cm.251130.csv", header = T)

# remove the huge CI secondary Nreabsorption, SD, and GR peaks:
all_QTL_CI <- all_QTL_CI %>% 
  filter(!chr==2 | !trait =="Nreabs") %>% 
  filter(!chr==6 | !trait =="SD") %>%
  filter(!chr==5 | !trait =="GR")

# assign a color to each trait
trait_colors <- c(
  SD = "#0072B2",
  GR = "#CC79A7",
  tough = "#8250C4",
  Nreabs = "#E69F00",
  Chlreabs = "#009E73"
)

# assign a label to each trait 
trait_labels <- c(
  SD        = "Seed dormancy",
  GR        = "Growth rate",
  tough     = "Leaf toughness",
  Nreabs    = "Nitrogen reabsorption during drought",
  Chlreabs  = "Chlorophyll reabsorption during drought"
)
# calculate average fst for each peak
# first calculate cm windows for fst data
fst_cm <- fst %>%
  arrange(chrN, cM_window_start) %>%
  group_by(chrN) %>%
  mutate(
    cM_window_end = lead(cM_window_start)
  ) %>%
  ungroup()

qtl_fst_cm <- all_QTL_CI %>%
  rowwise() %>%
  mutate(
    avg_fst_QTL = mean(
      fst_cm$avg_wc_fst[
        fst_cm$chrN == chr &
        fst_cm$cM_window_start <= upper95CI_cm &  # window starts before CI end
        fst_cm$cM_window_end   >= lower95CI_cm    # window ends after CI start
      ],
      na.rm = TRUE
    )
  ) %>%
  ungroup()

fst_windows_hist_w.QTL.CIs <- ggplot() +
  # Histogram
    geom_histogram(
      data = fst_summary,
      aes(x = avg_fst_per_cM, y = after_stat(density)),
      binwidth = 0.01,
    fill = "grey80",
    color = "black"
  ) +
  # QTL lines (legend ON + force point glyph)
  geom_vline(
    data = qtl_fst_cm,
    aes(xintercept = avg_fst_QTL, color = trait),
    linetype = "solid",
    size = 1,
    show.legend = TRUE,
    key_glyph = "point"     # <-- force legend to draw using point key
  ) +
  # Colors
  scale_color_manual(values = trait_colors, labels = trait_labels) +
  # Override legend aesthetics so points become squares
  guides(
    color = guide_legend(
      override.aes = list(
        shape = 15,     # square
        size = 5,
        linetype = 0    # no line in the legend
      )
    )
  ) +
  labs(
    x = "average differentiation (Fst in 1 cM windows)",
    y = "probability density",
    color = "QTL window averages\nTrait:"
  ) +
  theme_bw() +
  theme(
    legend.position = "right",
    legend.justification = "center"
  )


# save plot
# chr 2 SD peak is present, just under chr8 toughness peak
ggsave("fst_windows_hist_w.QTL.CIs.png", fst_windows_hist_w.QTL.CIs, device = "png", width = 17, height = 6.5, units = "cm")

# LEGEND note: Histogram of the probability density of average Fst in one cM windows across the genome. binwidth is 0.01, such that a y value of e.g. 5 indicates a 50% chance that a cM window average fst value will fall in that Fst bin range. 

### save both plots together ###
# Suppress legend for f_dM plot
p1 <- f_dM_windows_hist_w.QTL.CIs +
  guides(color = "none")

# Keep legend for fst plot
p2 <- fst_windows_hist_w.QTL.CIs

# Combine and collect only the legend from p2
f_dm.fst_hist_w.QTL.CIs <-
  p1 / p2 +
  plot_layout(guides = "collect")

ggsave("fst_windows_hist_w.6QTL.CIs.png", f_dm.fst_hist_w.QTL.CIs, device = "png", width = 13, height = 10, units = "cm")
```
